---
title: 'Econ 2367.02: Current Economic Issues in the United States'
subtitle: "WeFr 3:55PM - 5:15PM"
author: 'Instructor: Wei Yang Tham'
date: "Fall 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Email: tham.5\@osu.edu

Office: Arps 369

Office hours: Tuesday 2-3pm; Wednesday 2:30-3:30pm. Email me to set up an alternative time. 

# Course Overview

This is a topics class covering various economic issues and their impact on the economy and citizens of the United States. The aim of the course is to help you understand how to analyze policy issues using economic ways of thinking. You will also be developing your writing and critical thinking skills and learning how to formulate evidence based arguments. 

There is no textbook for this class. 

# Course objectives

## GE – Social Science Organizations and Polities 

Students understand the systematic study of human behavior and cognition; the structure of human societies, cultures, and institutions; and the processes by which individuals, groups, and societies interact, communicate, and use human, natural, and economic resources. 

1. Students understand the theories and methods of social scientific inquiry as they apply to the study of organizations and polities. 

2. Students understand the formation and durability of political, economic, and social organizing principles and their differences and similarities across contexts. 

3. Students comprehend and assess the nature and values of organizations and polities and their importance in social problem solving and policy making.  

## GE – Writing and Communication Level 2 

Students are skilled in written communication and expression, reading, critical thinking, oral expression, and visual expression.

1. Through critical analysis, discussion, and writing, students demonstrate the ability to read carefully and express ideas effectively. 

2. Students apply written, oral, and visual communication skills and conventions of academic discourse to the challenges of a specific discipline

# Grading

Assignment | Weight
---------- | ------
Quizzes    | 10%
Short essays | 20%
Research papers | 40%
Presentation | 20%
Participation | 10% 

## Readings
Readings will be announced at least one week ahead at the end of class. Readings on the reading list are subject to change as the course progresses. 

## Participation

Due to the nature of the course material, lectures will be discussion based. Students should come to class having carefully read the assigned materials and participate in the class discussion. If you exhibit a pattern of non-attendance and inattention during lectures and *especially* during student presentations your participation grade will suffer.

## Quizzes
Six times during the semester I will administer an in class quiz over the assigned readings that day. They are designed to incentivize reading, so they will be unannounced.  I will drop your lowest quiz grade.  Each quiz will be administered at the start of class.  Because I drop the lowest quiz, no make-ups will be administered.

## Short essays
Twice during the semester you will submit a 1-2 page **(not including works cited)** essay on an issue discussed (or to be discussed) in class. For each essay, you will be given prompts. You may choose the prompt you wish to respond to, but your position on the topic will be **randomly assigned**. Your job is to make the strongest possible case for your assigned position *regardless of your personal stance*, drawing on empirical evidence we have discussed in class as well as outside of class. 

## Research papers
You will be submitting more detailed research papers between 3 and 4 pages in length **not including works cited**.  It can be on any economic issue of your choosing. Your paper should have a strong thesis/hypothesis, which you will analyze and support by rigorous and cogent appeal to empirical evidence gathered and interpreted by you and others.   

## Paper Formatting 
All papers should be 1.5 spacing with 11 point Times New Roman font and 1-inch margins. Please include your name, paper title, and due date on a title page. All citations should follow Chicago author-date formatting rules (excluding the abstract), and an appropriate works cited page should ALWAYS be attached. You are, of course, encouraged to include any graphs or figures in your paper which you believe are helpful, but they do not count towards a page count, so please attach them at the end of the paper. Failure to follow these formatting instructions will result in the loss of 5 percent of the possible points for the assignment.

## Presentation
You will be presenting the findings from your final paper to the class. Presentations will last 10-15 minutes and you MUST respond to audience questions/criticisms. You will be allowed to bring up notes to refer to in response to questions, but the main body of the presentation must be done without notes. You must also have slides for the presentation. 

## Late Policy 
I do not accept late work. If you have a valid reason for needing an extension on an assignment please contact me in advance. Extenuating circumstances will be dealt with on a case-by-case basis.

## Final Grades

I round up fractional points *after* your final grade has been tallied i.e. a final score of 92.01 will get an A.

Letter Grade | Score
-------------| -----
A | 93
A-|90
B+|87
B|83
B-|80
C+|77
C|73
C-|70
D+|67
D|60

# Course schedule

This is the tentative schedule of topics. Changes will be announced in class. 

```{r}
library(tidyverse)
library(magrittr)
sched = read_csv("econ2367_schedule.csv") %>% select(Week, Date, Topics, Assignments)
sched %<>% replace_na(list(Assignments = ""))
knitr::kable(sched)
```

# Laptop and electronics policy

I recognize that laptop and tablets can have an important role to play in learning. To balance that against their potential to distract other students, I ask that you sit toward the back of the classroom if you are planning to use a laptop/tablet/other similar devices during class. This is not an attempt to penalize you, but to ensure that the classroom remains a conducive learning environment for all. 

# Resources

## Economics Learning Center
The Department of Economics runs the Economics Learning Center (ELC), where advanced undergraduates provide free tutoring for students in Economics 2001, 2002, 4001, and 4002. Assistance with other 3000-level and 4000-level classes is frequently available. The ELC is located in 311 Arps Hall (1945 North High Street) and is typically open from 9AM-5PM Monday-Thursday and 9AM-3PM on Fridays, opening each semester on the second (full) week of the term. Note that the purpose of the tutoring center is not to provide answers to assignments, but to help students learn economics. As the student, you are ultimately responsible for all course work you submit. 

## Ohio State Writing Center

> The Writing Center offers free help with writing at any stage of the writing process for any member of the university community. During our sessions, consultants can work with you on anything from research papers to lab reports, from dissertations to résumés, from proposals to application materials. Appointments are available in-person at 4120 Smith Lab, as well as for online sessions. You may schedule an in-person or online appointment by visiting WCOnline or by calling 614-688-4291. Please note that the Writing Center also offers daily walk-in hours—no appointment necessary—in Thompson Library. You do not have to bring in a piece of writing in order to schedule a writing center appointment. Many students report that some of their most productive sessions entail simply talking through ideas.   Please check out our Services [(https://cstw.osu.edu/writing-center/writing-center-services)](https://cstw.osu.edu/writing-center/writing-center-services)  page for the types of consultations we provide. We also maintain a Writing Resources [(https://cstw.osu.edu/writing-resources)]() page with writing handouts and links to online resources.

# Academic Misconduct

It is the responsibility of the Committee on Academic Misconduct to investigate or establish procedures for the investigation of all reported cases of student academic misconduct. The term "academic misconduct" includes all forms of student academic misconduct wherever committed; illustrated by, but not limited to, cases of plagiarism and dishonest practices in connection with examinations. Instructors shall report all instances of alleged academic misconduct to the committee (Faculty Rule 3335-5-487). For additional information, see the Code of Student Conduct [http://studentlife.osu.edu/csc/](http://studentlife.osu.edu/csc/).

# Student Life Disability Services

The University strives to make all learning experiences as accessible as possible. If you anticipate or experience academic barriers based on your disability (including mental health, chronic or temporary medical conditions), please let me know immediately so that we can privately discuss options. To establish reasonable accommodations, I may request that you register with Student Life Disability Services.  After registration, make arrangements with me as soon as possible to discuss your accommodations so that they may be implemented in a timely fashion. **SLDS contact information:** slds\@osu.edu; 614-292-3307; slds.osu.edu; 098 Baker Hall, 113 W. 12th Avenue.




