---
title: "Minimum Wage Teaching Notes"
author: "Tham, Wei Yang"
date: "7/13/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = TRUE)
```

## Facts about the minimum wage
```{r fed_minwage, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
library(magrittr)
fed_minwage = read_csv("minimum_wage/fed_minwage.csv")
cpi = read_csv("minimum_wage/cpi.csv")

minwage = inner_join(fed_minwage, cpi, by = "DATE") %>% 
  mutate(year = year(DATE)) %>% 
  rename(wage = FEDMINNFRWG, cpi2010 = CPALTT01USA661S)

minwage %<>% mutate(wage_real = wage/cpi2010)

minwage %>% gather(k, minwage, wage, wage_real) %>% 
  ggplot(aes(year, minwage, colour = k)) + 
  geom_line() + 
  labs(title = "in $2010")

```

- Federal minimum wage is $7.25 (since 2009)

## What does economic theory predict about raising about the minimum wage?
- Classical supply and demand predicts a disemployment effect (draw graph)
- (Dynamic) monopsony is consistent with no disemployment effect (draw graph)

## Empirical evidence
- Theory can only get us so far. What can we learn from actual changes in the minimum wage?
- (show meta-analysis graph)

## Card and Krueger
- asdf

## Explaining zero effect
- asdf

## What empirical research can and cannot tell us
- changing market conditions, institutions over time
- "local" changes in minimum wage

## More recent evidence 
- Jeff Clemens Great Recession paper
- Seattle 



 

