---
title: "Science Policy"
author: "Wei Yang Tham"
date: "September 27, 2017"
output:
  beamer_presentation:
    incremental: false
    theme: "metropolis"
    colortheme: "crane"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = 'pdf')
```

<!--
Learning objectives: 

Role of basic science in economic growth
economic rationale for public funding of science
how economics can be used to understand institutions and incentives in science - talk about doubling

-->

## Today

- Why do we think science and basic research are important?
- What is the economic justification for government funding of science?
- How do current scientific institutions incentive (or disincentivize) scientific discovery?

## Context
- For much of today's lecture, we will look at science policy through the lens of biomedical research
- The National Institutes of Health (NIH), the main source of government spending on biomedical research, is also the largest federal agency by R&D spending

## The Endless Frontier
- some history of US science?

## NIH Background
- NIH budget is over \$30 billion

## NIH funding has stagnated
```{r}
library(tidyverse)

nihfunding = read_tsv('nih_fundingfacts.tsv', col_types = cols(Statistic = 'd')) 
nihfunding %<>% select(-Source) %>% 
  rename(ic = IC, fiscalyear = `Fiscal Year`, topic = Topic, mechanism = Mechanism, activity = Activity, 
       type = Type, statistic = Statistic)

```

## 



