---
title: "Science Policy"
author: "Wei Yang Tham"
date: "September 27, 2017"
output:
  beamer_presentation:
    incremental: false
    theme: "metropolis"
    colortheme: "crane"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = 'pdf')
```

## Objectives

- Why do we think that science and basic research are important?
- What is the economic justification for government funding of science?
- How do current scientific institutions incentivize (or disincentivize) scientific discovery?

<!--
Learning objectives: 

Role of basic science in economic growth
economic rationale for public funding of science
how economics can be used to understand institutions and incentives in science - talk about doubling

podcast themes: much of it goes to university overhead; 10% chance of getting grant; what is the rate of return on science?

-->

## Quick recap of readings

1. Review of Mokyr's "Gifts of Athena"
    + What was the main reason suggested by Mokyr for why the Industrial Revolution was sustained beyond 1820?
    
2. Podcast on *The Wastebook*
    + Story of a scientist whose research (shrimp fight club) was used as an example of wasteful spending
    + Why is it so difficult to measure the returns on investment in basic research?
    
<!--
uncertain
hard to trace multiple uses
cumulativeness of knowledge: how far back do we go?
time horizon - uses we don't know of yet
-->

## Basic research is largely funded by the government
```{r, fig.align='center'}
library(tidyverse)
library(magrittr)
theme_set(theme_minimal(base_size = 22))


basic = read_tsv("basic_bysource.tsv")
years = sort(rep(1953:2015, 4))

basic %>% 
  mutate(Year = years) %>% 
  ggplot(aes(Year, Funding/1000, colour = `Funding Source`)) + 
  geom_line(size = 2) + 
  labs(y = "Billions", x = "Fiscal Year")

```

## Group Discussion
- From groups of 3-4
- In your groups, come up with a few reasons for why you think government should and should not invest in basic research
    + In your discussion, think about the economic arguments typically used to justify government subsidization or provision of certain goods
    
## Some arguments for government funding of basic research
- Knowledge is a public good; hard to exclude others (non-excludable) and it is not depleted when shared (non-rivalrous)
- Therefore market will under-provide basic research
- Risky
- Benefits may take a long time to be realized (if at all)

## Some arguments against government funding of basic research
- "Government" is bad at picking winners and losers
- Government funding crowds out private investment, so there is no net benefit

## Biomedical research in the US
- For much of today's lecture, we will look at science policy through the lens of biomedical research
- The National Institutes of Health (NIH), the main source of government spending on biomedical research, is also the largest federal agency by R&D spending

## NIH Background
- The NIH annual budget is over \$30 billion
- A collection of 27 Institutes and Centers, each focused on different diseases or areas of science (e.g. National Cancer Institute, National Heart, Lung and Blood Institute, National Institute of Aging)

## How does the NIH decide who gets funding?
- Scientists submit applications for grants
- Applications are reviewed and scored by panels of other scientists that work in a similar area of research
- Note that scientists do not themselves decide which applications are approved (there is a fairly complicated process that does this *after* applications have been scored)

## NIH funding has fallen in real terms
```{r}

nihfunding = read_tsv("nih_fundingfacts.tsv", col_types = cols(Statistic = 'd')) 
nihfunding %<>% select(-Source) %>% 
  rename(ic = IC, fiscalyear = `Fiscal Year`, topic = Topic, mechanism = Mechanism, activity = Activity, 
       type = Type, statistic = Statistic)

nihfunding %<>% spread(topic, statistic) %>% 
  rename(totalfunding = `Awards - Funding (Total Cost)`, meanfunding = `Awards - Average Total Cost`, 
         n_awards = `Awards - Number`, success_rate = `Success Rate`, 
         n_applications = `Applications - Number`)

brdpi = read_tsv('brdpi.tsv')
# Base year 1998
brdpi %<>% mutate(brdpi98 = max(if_else(fiscalyear == 1998, brdpi, NA_real_), na.rm = T), 
                  brdpi = brdpi/brdpi98)

nihfunding %>% 
  left_join(brdpi, by = 'fiscalyear') %>% 
  filter(ic == "All NIH", mechanism == 'Research Grants', activity == 'R01', 
         type == 'All Types (in aggregate)')  %>% 
  ggplot(aes(fiscalyear, totalfunding/brdpi/1e9)) + 
  geom_line(size = 1.2) + geom_point(size = 3) +  
  geom_vline(xintercept = 2003, linetype = 2) +
  geom_label(data = data_frame(fiscalyear = 2005.3, y = 6.5, text = "Doubling Ends"), 
            aes(x = fiscalyear, y = y, label = text), size = 4.5, fill = 'darkred', color = 'white') +
  labs(title = "Total funding for R01 grants", 
       subtitle = "Deflated by Biomedical R&D Price Index (Base Year 1998)", 
       y = "$ Billion", x = "Fiscal Year") 

```

## Chances of getting a grant have fallen drastically
```{r}
nihfunding %>% 
  filter(ic == "All NIH", mechanism == 'Research Grants', activity == 'R01', 
         type == 'All Types (in aggregate)')  %>% 
  ggplot(aes(fiscalyear, success_rate*100)) + 
  geom_line(size = 1.2) + geom_point(size = 3) + 
  geom_vline(xintercept = 2003, linetype = 2) + 
  geom_label(data = data_frame(fiscalyear = 2005.3, success_rate = 0.3, text = "Doubling Ends"), 
            aes(label = text), size = 4.5, fill = 'darkred', color = 'white') + 
  labs(x = "Fiscal Year", y = "Percentage points")
```

## The Shopping Mall model of science
- Recall in the podcast that the scientist describes how substantial amount of money goes to overhead
- Shopping mall model: Universities build facilities and "rent" them to faculty
- This "rent" is known as an "indirect cost"
- Universities usually provide some startup funds; after that faculty have to get their own funding
- Some faculty are on "soft money" positions i.e. they have to raise their own salaries

## Question
- Recall how 


