---
title: "Science Policy"
author: "Wei Yang Tham"
date: "September 27, 2017"
output:
  beamer_presentation:
    incremental: true
    theme: "metropolis"
    colortheme: "crane"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = 'pdf')
```

## Objectives

- Why do we think that science and basic research are important?
- What is the economic justification for government funding of science? Does the empirical evidence support this?
- How do current scientific institutions incentivize (or disincentivize) scientific discovery?

<!--
Learning objectives: 

Role of basic science in economic growth
economic rationale for public funding of science
how economics can be used to understand institutions and incentives in science - talk about doubling

podcast themes: much of it goes to university overhead; 10% chance of getting grant; what is the rate of return on science?

-->

## Quick recap of readings

1. Review of Mokyr's "Gifts of Athena"
    + What was the main reason suggested by Mokyr for why the Industrial Revolution was sustained beyond 1820?
    
2. *The Wastebook* podcast
    + Story of a scientist whose research ("shrimp fight club") was used as an example of wasteful spending
    + Why is it so difficult to measure the returns on investment in basic research?
    
<!--
uncertain
hard to trace multiple uses
cumulativeness of knowledge: how far back do we go?
time horizon - uses we don't know of yet
-->

## Basic research is largely funded by the government
```{r, fig.align='center'}
library(tidyverse)
library(magrittr)
theme_set(theme_minimal(base_size = 22))


basic = read_tsv("basic_bysource.tsv")
years = sort(rep(1953:2015, 4))

basic %>% 
  mutate(Year = years) %>% 
  ggplot(aes(Year, Funding/1000, colour = `Funding Source`)) + 
  geom_line(size = 2) + 
  labs(y = "Billions", x = "Fiscal Year")

```

## Group Discussion
- From groups of 3-4
- In your groups, come up with a few reasons for why you think government should and should not invest in basic research
    + In your discussion, think about the economic arguments typically used to justify government subsidization or provision of certain goods
    
## Some arguments for government funding of basic research
- Knowledge is a public good; hard to exclude others (non-excludable) and it is not depleted when shared (non-rivalrous)
- Therefore market will under-provide basic research
- Risky
- Benefits may take a long time to be realized (if at all)

## Some arguments against government funding of basic research
- "Government" is bad at picking winners and losers
- Government funding crowds out private investment, so there is no net benefit

## Biomedical research in the US
- For much of today's lecture, we will look at science policy through the lens of biomedical research
- The National Institutes of Health (NIH) is the largest federal agency by funding of basic research

## NIH Background
- The NIH annual budget is over \$30 billion
- A collection of 27 Institutes and Centers, each focused on different diseases or areas of science (e.g. National Cancer Institute, National Heart, Lung and Blood Institute, National Institute of Aging)
- Besides research grants, the NIH also provides funds for training and research facilities

## How does the NIH decide who gets grants?
- Scientists submit applications for grants
- Applications are reviewed and scored by panels of other scientists that work in a similar area of research
- Note that scientists do not themselves decide which applications are approved (there is a fairly complicated process that does this *after* applications have been scored)

## The Shopping Mall model of science
- Recall from the podcast that the scientist describes how a substantial amount of her grant money goes to overhead
- Shopping mall model: Universities build facilities and "rent" them to faculty
- This "rent" is known as an "indirect cost"
- Universities usually provide some startup funds; after that faculty have to get their own funding
- Some faculty are on "soft money" positions i.e. they have to raise their own salaries

## NIH funding has fallen in real terms
```{r}

nihfunding = read_tsv("nih_fundingfacts.tsv", col_types = cols(Statistic = 'd')) 
nihfunding %<>% select(-Source) %>% 
  rename(ic = IC, fiscalyear = `Fiscal Year`, topic = Topic, mechanism = Mechanism, activity = Activity, 
       type = Type, statistic = Statistic)

nihfunding %<>% spread(topic, statistic) %>% 
  rename(totalfunding = `Awards - Funding (Total Cost)`, meanfunding = `Awards - Average Total Cost`, 
         n_awards = `Awards - Number`, success_rate = `Success Rate`, 
         n_applications = `Applications - Number`)

brdpi = read_tsv('brdpi.tsv')
# Base year 1998
brdpi %<>% mutate(brdpi98 = max(if_else(fiscalyear == 1998, brdpi, NA_real_), na.rm = T), 
                  brdpi = brdpi/brdpi98)

nihfunding %>% 
  left_join(brdpi, by = 'fiscalyear') %>% 
  filter(ic == "All NIH", mechanism == 'Research Grants', activity == 'R01', 
         type == 'All Types (in aggregate)')  %>% 
  ggplot(aes(fiscalyear, totalfunding/brdpi/1e9)) + 
  geom_line(size = 1.2) + geom_point(size = 3) +  
  geom_vline(xintercept = 2003, linetype = 2) +
  geom_label(data = data_frame(fiscalyear = 2005.3, y = 6.5, text = "Doubling Ends"), 
            aes(x = fiscalyear, y = y, label = text), size = 4.5, fill = 'darkred', color = 'white') +
  labs(title = "Total funding for R01 grants", 
       subtitle = "Deflated by Biomedical R&D Price Index (Base Year 1998)", 
       y = "$ Billion", x = "Fiscal Year") 

```

## Chances of getting a grant have fallen drastically
```{r}
nihfunding %>% 
  filter(ic == "All NIH", mechanism == 'Research Grants', activity == 'R01', 
         type == 'All Types (in aggregate)')  %>% 
  ggplot(aes(fiscalyear, success_rate*100)) + 
  geom_line(size = 1.2) + geom_point(size = 3) + 
  geom_vline(xintercept = 2003, linetype = 2) + 
  geom_label(data = data_frame(fiscalyear = 2005.3, success_rate = 0.3, text = "Doubling Ends"), 
            aes(label = text), size = 4.5, fill = 'darkred', color = 'white') + 
  labs(x = "Fiscal Year", y = "Percentage points")
```

## Question
- One of the justifications for government funding of basic research is that it is better able to bear the riskiness of basic research
- How might the shopping mall model work against that goal, especially in the current (biomedical) funding environment? Hint: consider the scientist's point of view and how limited funding might affect the type of projects she pursues

## Is more funding the solution?
- Imagine that as a policymaker you had \$30bn to add to the NIH budget. You could do this either by allocating it all to next year's budget, or spread it out over the next 10 years. 
- Quick survey: Which option do you think is better?

## What If Congress Doubled R&D Spending? (Freeman and Van Reenen 2009)
- From 1998 to 2003, the NIH budget was increased every year until it had doubled (nominally) in 2003
- In 2004, this rapid increase was brought to a sudden and unexpected halt

## Nominal funding has stagnated since 2003
```{r}
nihfunding %>% 
  filter(ic == "All NIH", mechanism == 'Research Grants', activity == 'R01', 
         type == 'All Types (in aggregate)')  %>% 
  ggplot(aes(fiscalyear, totalfunding/1e9)) + 
  geom_line(size = 1.2) + geom_point(size = 3) +  
  geom_vline(xintercept = 2003, linetype = 2) +
  geom_label(data = data_frame(fiscalyear = 2005.3, y = 6.5, text = "Doubling Ends"), 
            aes(x = fiscalyear, y = y, label = text), size = 4.5, fill = 'darkred', color = 'white') +
  labs(title = "Total funding for R01 grants", 
       subtitle = "Nominal", 
       y = "$ Billion", x = "Fiscal Year") 
```

## Discussion
- Why might a sudden increase in funding not lead to an increase in scientific output in the short-term?
- It might help you to think of a research lab as a firm. What is its production function? What are its inputs? How would the markets for its inputs respond to an increase in demand?

<!--
market for PIs inelastic
market for grad students more elastic
lab expansion takes time; for a major shock like this may need new buildings
-->

## Research lab production function
- Output: knowledge, often measured by papers, citations, text analysis
- Labor inputs: principal investigator (the boss), postdocs, gradute students, staff scientists
- Capital inputs: equipment (e.g. MRI machine, test tubes), animals

## Consequences of the doubling 
- Essentially led to a biomedical science "bubble"
- Universities invested heavily in new buildings and increased hiring, leading to excess capacity and layoffs as funding stagnated
- Rise in hiring of postdocs and graduate students, but with poor prospects for employment as researchers due to dropoff in funding after 2003

## Concerns in the post-doubling era

> "<span style="color:black">The system now **favors those who can guarantee results** </span>rather than those with potentially path-breaking ideas that, by definition, cannot promise success." - Alberts et al (2014)

## If "how much" is not the answer, what about "how"?
- Awarded for 3-5 years
- When grant expires, have to apply for a renewal
- Typical NIH grant is evaluated based on the *project*
- What if grants were awarded to the *person*?

## Question
- Why might awarding grants to the person rather than the project be more effective in generating breakthrough science?

## HHMI awards
- Funds "people, not projects"
- Prestigious; intended for scientists with high potential
- HHMI investigators appointed for 5 years initially, usually renewed at least once
- Receive high-quality feedback at first review
- First review more focused on whether ideas have potential to be high impact rather than accomplishments

## Azoulay, Graff-Zivin, Manso (2010)
- Compare HHMI winners to other high-achieving scientists
- Find that HHMI winners are more likely to explore new lines of research and generate high-impact research

## Discussion
- Why might we not want the NIH to adopt the HHMI model for all of its grants?

<!--
want some incremental science as well as radical science
time-intensive for reviewers
HHMI population is most talented 
-->

## What is the impact of NIH funding?
- We've seen that the biomedical research system could be improved
- But how well is it doing right now?

## Two studies
- We'll look at two studies that approach the problem differently
- One looks at the impact of getting NIH funding on the individual researchers
- The other looks at the impact of getting NIH funding on a scientific area

## Jacob and Lefgren (2007)
- "The Impact of Research Grant Funding on Scientific Productivity"
- Regression discontinuity strategy: compare the publication output of researchers who were just above the threshold for getting a grant to those just below the threshold
- Find that getting NIH funding results in only a small or zero increase in number of publications 

## Discussion
- Does this necessarily imply that the impact of NIH funding on research is zero? 
    + Hint: Imagine you're an entrepreneur trying to get some startup capital and you got turned down by a potential investor. What's a natural next step?
    
<!--
marginal grant
substitute with other funds
-->
    
## Other considerations
- Are publications a good outcome measure?
- Recall the study on the H-1B visa lottery
    + Not so informative about aggregate changes in funding
    
    
## Azoulay, Graff-Zivin, Li, and Sampat (2015)
- "Public R&D Investments and Private-Sector Patenting: Evidence from NIH Funding Rules"
- Look at the impact of NIH funding on research areas rather than individuals
- Instrumental variables approach
- Look at effect of NIH funding on *total* patent output in a research area to account for crowd-out
    + If there is complete crowding-out, then total patent output should be unchanged

## Results
- Estimate that a \$10 million increase in funding leads to 2.3 new patents
- \$1 in NIH funding leads to \$1.40 in drug sales, not including other benefits like development of devices or public health improvements

## Summary
- Knowledge is a public good
- Economic theory suggests that the production of knowledge, especially through basic research, should be subsidized by the government
- The doubling of the NIH budget from 1998-2003 showed that increasing science funding, if not done carefully, can lead to undesirable consequences
- Jacob and Lefgren (2007) suggests that on the margins, researchers are able to find other ways to fund their research if they do not get NIH funding
- Azoulay et al (2015) find that research areas which receive more NIH funding lead to more follow-on innovation in the form of patents

## Readings

**Wednesday**

[\*TED Talk by David Autor](https://www.ted.com/talks/david_autor_why_are_there_still_so_many_jobs)

**Friday**

[\**The Basic Income experiment*, Planet Money podcast](http://www.npr.org/sections/money/2017/09/22/552850245/episode-796-the-basic-income-experiment)

[\**Why Some People Love Tax Day*, Planet Money podcast](http://www.npr.org/sections/money/2013/04/12/177063399/episode-451-why-some-people-love-tax-day)

[\*Summary of evidence on cash transfers in developing countries, GiveDirectly website](https://www.givedirectly.org/research-on-cash-transfers)




